Considering scale in multilayer animal social networks

### 2020-06-23 paper/manuscript.Rmd
minor edits discussion

Multilayer networks represent a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, for example as defined by habitat type as in
Figure \@ref(fig:lcres). Since habitat data are available across a range of
spatial scales, we evaluated the influence of increasing spatial resolution on
multilayer network metrics. As spatial resolution increased, edge overlap
remained stable in open and forest layers but decreased in foraging layers
(Figure \@ref(fig:socres) D). Graph strength did not change in forest layers
with increasing spatial resolution (Figure \@ref(fig:socres) E). Foraging and
open layer had greater individual variation in graph strength, and foraging
layers showed an increase in graph strength with increasing spatial resolution
(Figure \@ref(fig:socres) E). Foraging habitat is rare (14.5% of the landscape)
on Fogo Island and the proportion of relocations in foraging habitat decreased
as spatial resolution increased (14.5% at 30 m - <2% at 1000 m). Despite this,
at 30 m resolution the foraging layer had similar edge overlap as the open layer
indicating its importance in caribou sociality. As such, multilayer network
measures related to habitat are explicitly dependent on spatial scale.  If the
spatial resolution of the habitat data are insufficient to capture rare
habitats, for example, context-specific relationships between individuals may
not be captured.
<!-- habitat+social -->

@@

<!-- temp --> 
Changes in network metrics through time can explain the influence
of seasonality of resources on animal sociality. Using an ordinal multilayer
network composed of 20 discrete time windows throughout the study period, we
observed periods of increased sociality in caribou on Fogo Island. Edge overlap
was higher during fall, winter and spring (approximately October - June) than in
summer (approximately June - October)  (Figure \@ref(fig:temp) A). Similarly,
individual graph strength was also highest in these periods (Figure
\@ref(fig:temp) C). Periods of increased sociality correspond to seasons of
heterogeneous resource availability [Peigner_2019]. By explicitly considering
the temporal context of social interactions in separate layers, multilayer
networks provide a detailed and dynamic perspective of animal social networks.
In addition, this could be extended using time-aggregated networks and a
moving-window approach to inform the temporal scale including time window size
and sampling effort [@Farine_2017; @Bonnell_2020]. 
<!-- temp -->

<!-- nobs -->
It is crucial to include sufficient data within sampling periods or time windows
to construct robust networks [@Farine_2017; Proskurnikov_2017]. To test the
influence of data quantity on multilayer networks, we used a subset of
observations to generate multilayer networks. Variance decreased across all
network metrics in all three habitat layers with increasing number of
observations. Individual graph strength showed high variability at low number of
observations, while edge overlap and multidegree increased sharply. After
approximately 200 observations, the rate of change for all metrics decreased and
and graph strength remained relatively static, while edge overlap and
multidegree continued to increase slowly. As more observations were added, rarer
relationships were recorded revealing the value of this form of sensitivity
analysis. Finally, due to the autocorrelated nature of biologging data it is
essential to ensure social association data area non-randomly distributed
[@Farine_2017b].
<!-- nobs -->

<!-- (this was the) conclusion -->

Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. Recent technological advances in
remotely sensed imagery [@Drusch_2012] and animal tracking devices
[@B_rger_2016; @Wikelski_2007] have increased the availability of complex, high
resolution data. We highlight ways in which scale should be considered when
building social networks from both observed and remotely-sensed data. It is
important to consider a broad continuum of scales to fully capture ecological
processes, carefully select and combine sources of data with different temporal
and spatial scales, and establish a clear link between scales of observation and
meaningful biology.

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. We demonstrate the novel analyses in varying social, spatial,
and temporal scale within a multilayer framework. We show that (1) multilayer
social networks are scale-dependent, and (2) the use of multilayer networks is
useful for identifying habitat or season-specific social processes. Our analysis
of caribou multilayer networks incorporating high resolution GPS telemetry data
and spatial land cover data is grounded in the context of observed animal social
networks, as the scales we chose to explore, e.g., number of observations and
sociality across space and time, are ones that have also traditionally been done
in ethology.

<!-- TODO: round it out -->

<!-- (this was the) conclusion -->

>>>>

Multilayer networks represent a novel framework for explicitly testing the
influence of habitat selection on sociality across scales. Animal social systems
can be parsed by spatial contexts, for example as defined by habitat type as in
Figure \@ref(fig:lcres). Since habitat data are available across a range of
spatial scales, we evaluated the influence of increasing spatial resolution on
multilayer network metrics. As spatial resolution increased, edge overlap
remained stable in open and forest layers but decreased in foraging layers
(Figure \@ref(fig:lcres) D). Graph strength did not change in forest layers
with increasing spatial resolution (Figure \@ref(fig:lcres) E). Foraging and
open layer had greater individual variation in graph strength, and foraging
layers showed an increase in graph strength with increasing spatial resolution
(Figure \@ref(fig:lcres) E). Foraging habitat is rare (12.8% of the landscape)
on Fogo Island and the proportion of relocations in foraging habitat decreased
as spatial resolution increased (14.5% at 30 m - <2% at 1000 m). Despite this,
the foraging layer had similar edge overlap as the open layer at 30 m resolution 
indicating its importance in caribou sociality. As such, multilayer network
measures related to habitat are explicitly dependent on spatial scale.  If the
spatial resolution of the habitat data are insufficient to capture rare
habitats, for example, context-specific relationships between individuals may
not be captured.
<!-- habitat+social -->

@@

<!-- temp --> 
Changes in network metrics through time can explain the influence
of seasonality of resources on animal sociality. Using an ordinal multilayer
network composed of 20 discrete time windows throughout the study period, we
observed periods of increased sociality in caribou on Fogo Island. Edge overlap
was higher during fall, winter and spring (approximately October - June) than in
summer (approximately June - October)  (Figure \@ref(fig:temp) A). Similarly,
individual graph strength was also higher in these periods (Figure
\@ref(fig:temp) C). These periods of increased sociality correspond to seasons of
heterogeneous resource availability [Peigner_2019]. By explicitly considering
the temporal context of social interactions in separate layers, multilayer
networks provide a detailed and dynamic perspective of animal social networks.
In addition, this could be extended using time-aggregated networks and a
moving-window approach to inform the temporal scale including time window size
and sampling effort [@Farine_2017; @Bonnell_2020]. 
<!-- temp -->

<!-- nobs -->
It is crucial to include sufficient data within sampling periods or time windows
to construct robust networks [@Farine_2017; Proskurnikov_2017]. To test the
influence of data quantity on multilayer networks, we generate multilayer networks
using a subset of observations. Variance decreased across all
network metrics in all three habitat layers with increasing number of
observations. Individual graph strength showed high variability at low number of
observations, while edge overlap and multidegree increased sharply. After
approximately 200 observations, the rate of change for all metrics decreased and
and graph strength remained relatively static, while edge overlap and
multidegree continued to increase slowly. As more observations were added, rarer
relationships were recorded revealing the value of this form of sensitivity
analysis. Finally, due to the autocorrelated nature of biologging data it is
essential to ensure social association data area non-randomly distributed
[@Farine_2017b].
<!-- nobs -->

<!-- conclusion -->
Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. Recent technological advances in
remotely sensed imagery [@Drusch_2012] and animal tracking devices
[@B_rger_2016; @Wikelski_2007] have increased the availability of complex, high
resolution data. We highlight ways in which scale should be considered when
building social networks from both observed and remotely-sensed data. It is
important to consider a broad continuum of scales to fully capture ecological
processes, carefully select and combine sources of data with different temporal
and spatial scales, and establish a clear link between scales of observation and
meaningful biology.

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. We demonstrate the novel analyses in varying social, spatial,
and temporal scale within a multilayer framework. We show that (1) multilayer
social networks are scale-dependent, and (2) the use of multilayer networks is
useful for identifying habitat or season-specific social processes. Our analysis
of caribou multilayer networks incorporating high resolution GPS telemetry data
and spatial land cover data is grounded in the context of observed animal social
networks, as the scales we chose to explore, e.g., number of observations and
sociality across space and time, are ones that have also traditionally been done
in ethology.
<!-- TODO: round it out -->

<!-- conclusion -->
